---
title: "Experiment Data Processing Report for DBAT_1_2021"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: true
    number_sections: true
params: 
  topography: false
  ssurgo: false
  ol_yield_sd_factor: 4
  overlap_acceptance_pct: 0.1
---

<style type="text/css">

body{ /* Normal  */
      font-size: 20px;
  }
td {  /* Table  */
  font-size: 16px;
}
h1.title {
  font-size: 56px;
}
h1 { /* Header 1 */
  font-size: 48px;
}
h2 { /* Header 2 */
    font-size: 36px;
}
h3 { /* Header 3 */
  font-size: 24px;
}
code.r{ /* Code block */
    font-size: 16px;
}
</style>

```{r setup, cache = F, echo = F}
library(knitr)
knitr::opts_chunk$set(
  cache = FALSE,
  echo = TRUE,
  error = TRUE,
  warning = FALSE,
  cache.lazy = FALSE,
  message = FALSE,
  fig.width = 12,
  fig.height = 12
)
```

```{r packages, cache = FALSE}
# === packages ===#
library(raster)
library(sf)
library(here)
library(ggplot2)
library(tmap)
library(patchwork)
library(measurements)
library(scam)
library(stars)
library(parallel)
library(future.apply)
library(dplyr)
library(tidyverse)
library(modelsummary)
library(jsonlite)
library(data.table)
library(exactextractr)
library(future.apply)
```

```{r preparations, cache = F}
source(here::here("Functions/prepare.R"))

#--- github ---#
source("https://raw.githubusercontent.com/brittanikedge/DIFM/main/Functions.R")

#--- define field-year ---#
ffy <- "DBAT_1_2021"

#--- read the variable name dictionary ---#
dictionary <-
  fromJSON(
    file.path(
      here("Data", "CommonData"),
      "variable_name_dictionary.json"
    ),
    flatten = TRUE
  ) %>%
  data.table()

#--- define field parameters ---#
trial_info <- get_trial_parameter(ffy)
input_trial_data <- trial_info$input_data_trial

#---  ---#
plan(multiprocess, workers = detectCores() - 2)
```

```{r set-root-dir}
opts_knit$set(root.dir = here("Data", "Growers", ffy))
```

```{r map-layout, cache = TRUE}
tm_layout_to_add <-
  tm_layout(
    legend.outside = "TRUE",
    frame = FALSE,
    legend.title.size = 2,
    legend.text.size = 1.5
  )
```

# Yield Data Processing

## Preparation

### Read datasets

```{r Reading-yield-data}
#--- yield ---#
yield <-
  here(
    "Data/Growers",
    ffy,
    "Raw/21_DBAT_C_ML_HV.shp"
  ) %>%
  st_read() %>%
  # === this took effect only when crs is missing ===#
  st_set_4326() %>%
  st_transform_utm() %>%
  setnames(names(.), tolower(names(.)))
```

```{r viz-yield-data}
tm_shape(yield) +
  tm_dots() +
  tm_layout_to_add
```

## Various checks and consistency

### Change variable names

```{r Changing-variable-names}
if ("time" %in% names(yield) & "isotime" %in% names(yield)) {
  yield <- dplyr::select(yield, -time)
}

dict_yield <- dictionary[type == "yield", ]
col_list <- dict_yield[, column]

yield <- make_var_name_consistent(
  yield,
  dict_yield
)

yield <- dplyr::select(yield, matches(col_list)) %>%
  mutate(yield_id := 1:nrow(.)) %>%
  filter(!is.na(yield_vol))

if ("offset" %in% names(yield) == FALSE) {
  yield$offset <- 0
}
```

### Order by time

```{r eval = F}
setorder(yield, time)
```

### Check units

This code assumes that if `distance` is in meter, then other variables (e.g., `width`) are also measured in meter.

```{r Checking-units-of-measurements}

dist_test <-
  yield %>%
  dplyr::select(distance) %>%
  cbind(., st_coordinates(.)) %>%
  data.table() %>%
  .[, dif_X := c(0, diff(X))] %>%
  .[, dif_Y := c(0, diff(Y))] %>%
  #--- distance in meter ---#
  .[, distance_r := sqrt(dif_X^2 + dif_Y^2)] %>%
  #--- dif in difference in meter and recorded distance ---#
  .[, dif_distance := abs(distance - distance_r)] %>%
  .[, dif_distance_conv := abs(conv_unit(distance, "ft", "m") - distance_r)]

# we want to see if the converted units from ft to meter are closer to the distance measured in r
if (mean(dist_test$dif_distance_conv, na.rm = TRUE) < mean(dist_test$dif_distance, na.rm = TRUE)) {
  units <- "imperial"
} else {
  units <- "metric"
}
```

### transform measurement unit

```{r Converting-units-to-metric}

if (units == "imperial") {
  yield <- yield %>%
    mutate(
      width = conv_unit(width, "ft", "m"),
      distance = conv_unit(distance, "ft", "m"),
      offset = conv_unit(offset, "ft", "m")
    )
} else {
  yield <- yield
}
```

### save the yield point data

```{r }
saveRDS(
  yield,
  here(
    "Data/Growers",
    ffy,
    "Intermediate/yield_point.rds"
  )
)
```

## Group points by angle

```{r Grouping-yield-points-by-angle-and-detecting-sharp-turns-and-changes-of-rows}

yield_grouped <- group_points_sc(yield, by_var = NA, 30)
```

Red points are the points that are recognized for making a sharp turn from their respective previous points.

```{r }
tm_shape(yield_grouped) +
  tm_dots(col = "blue") +
  tm_shape(filter(yield_grouped, change_group == TRUE)) +
  tm_dots(col = "red", size = 0.1) +
  tm_layout_to_add
```

## Intersect with the Trial Design and Flag Outliers by Treatment Block

### Import the trial design and visualize

**Read the trial design shape file**

```{r }
trd_polygons <-
  here(
    "Data/Growers", ffy, "TrialDesign",
    paste0(trial_info$tr_design_data, ".shp")
  ) %>%
  st_read() %>%
  mutate(treat_id = 1:nrow(.)) %>%
  dplyr::rename_at(vars(-one_of("geometry")), ~ paste0("trd_", .))

if (!("trd_PlotNum" %in% names(trd_polygons))) {
  trd_polygons <- mutate(trd_polygons, trd_PlotNum = 1:nrow(trd_polygons))
}
```

**Visualization**
```{r }
# === vis ===#
tm_shape(trd_polygons) +
  tm_polygons(col = "trd_Trt")
```

### Flag yield outliers by treatment block

```{r }

params <- list()
params$ol_yield_sd_factor <- 4

sd_factor <- params$ol_yield_sd_factor

yield_trd <-
  yield_grouped %>%
  dplyr::select(yield_id, yield_vol, geometry) %>%
  #* identify which yield point falls in which treatment block
  st_join(., trd_polygons) %>%
  #* order by yield id
  arrange(yield_id) %>%
  data.table() %>%
  #* flag outliers by treatment block
  nest_by(trd_PlotNum) %>%
  mutate(data = list(
    flag_bad_points(
      data = data,
      var_name = "yield_vol",
      sd_factor = sd_factor
    )
  )) %>%
  unnest() %>%
  dplyr::select(-geometry, -yield_vol) %>%
  data.table()

#* join with yield_grouped
yield_grouped <-
  left_join(yield_grouped, yield_trd, by = "yield_id")
```

**Show the yield points that are flagged as outliers**

1 indicates outliers.

```{r }
yield_grouped <- mutate(yield_grouped, ol_yield_vol_txt = ifelse(ol_yield_vol == 1, "Yes", "No"))

g_y_outlier <-
  ggplot(yield_grouped) +
  geom_sf(aes(color = ol_yield_vol_txt), size = 0.3) +
  scale_color_manual(
    values = c("Yes" = "red", "No" = "blue"),
    name = "Outlier in Yield Value"
  ) +
  guides(color = guide_legend(override.aes = list(size = 3))) +
  theme_void() +
  theme(
    legend.position = "bottom"
  )
```


### Flag speed anomalies

```{r Flagging-outliers}
# /*----------------------------------*/
#' ## Speed-based
# /*----------------------------------*/
is_speed_available <- !all(is.na(yield_grouped$speed))

if (is_speed_available) {
  yield_grouped <-
    flag_bad_points(
      yield_grouped,
      "speed",
      sd_factor = sd_factor
    )
} else {
  yield_grouped <-
    yield_grouped %>%
    mutate(ol_speed = 0)
}
```

**Yield points flagged based on `speed`**

```{r }
yield_grouped <- mutate(yield_grouped, ol_speed_txt = ifelse(ol_speed == 1, "Yes", "No"))

g_s_outlier <-
  ggplot(yield_grouped) +
  geom_sf(aes(color = ol_speed_txt), size = 0.3) +
  scale_color_manual(
    values = c("Yes" = "red", "No" = "blue"),
    name = "Outlier in Driving Speed"
  ) +
  guides(color = guide_legend(override.aes = list(size = 3))) +
  theme_void() +
  theme(
    legend.position = "bottom"
  )
```

```{r}
g_outliers <- g_y_outlier | g_s_outlier

ggsave(here("Misc/g_outliers.pdf"), g_outliers, width = 7, height = 4)
ggsave(here("Misc/g_outliers.png"), g_outliers, width = 7, height = 4, dpi = 600)
```

### Identify yield points that are close to either the yield and speed outliers

**Identify points that are close to outliers**
```{r }
close_to_ol_id <-
  yield_grouped %>%
  nest_by(group, trd_PlotNum) %>%
  mutate(outliers = list(
    #* keep those that are either yield or speed outliers
    filter(data, ol_yield_vol == 1 | ol_speed == 1) %>%
      dplyr::select(ol_yield_vol) %>%
      st_buffer(dist = 10)
  )) %>%
  mutate(close_to_outliers = list(
    filter(data, ol_yield_vol == 0 & ol_speed == 0) %>%
      dplyr::select(yield_id) %>%
      st_join(., outliers) %>%
      filter(!is.na(ol_yield_vol)) %>%
      pull(yield_id) %>%
      unique()
  )) %>%
  pull(close_to_outliers) %>%
  reduce(c) %>%
  unique()
```

**Create a variable that takes 1 if either an outlier or close-to-outliers**
```{r }
yield_grouped <-
  yield_grouped %>%
  mutate(
    drop_as_outliers =
      ifelse(
        yield_id %in% c(close_to_ol_id) | ol_yield_vol == 1 | ol_speed == 1,
        1, 0
      )
  )
```

**Visualize the data points that are flagged to be dropped **

Red points are flagged as either outliers themselves or close to outliers
```{r }
yield_grouped <- mutate(yield_grouped, drop_as_outliers_txt = ifelse(drop_as_outliers == 1, "Yes", "No"))

g_outliers_all <-
  ggplot(yield_grouped) +
  geom_sf(aes(color = drop_as_outliers_txt), size = 0.3) +
  scale_color_manual(
    values = c("Yes" = "red", "No" = "blue"),
    name = "Dropped as Outliers"
  ) +
  guides(color = guide_legend(override.aes = list(size = 3))) +
  theme_void() +
  theme(
    legend.position = "bottom"
  )

ggsave(here("Misc/g_outliers_all.pdf"), g_outliers_all, width = 7, height = 4)
ggsave(here("Misc/g_outliers_all.png"), g_outliers_all, width = 7, height = 4, dpi = 600)
```


### Yield histogram before and after dropping flagged observations

```{r }
dt_before <-
  data.table(yield_grouped) %>%
  .[, .(yield_vol)] %>%
  .[, type := "Before"]

dt_after <-
  data.table(yield_grouped) %>%
  .[drop_as_outliers == 0, ] %>%
  .[, .(yield_vol)] %>%
  .[, type := "After"]

dt_dropped <-
  data.table(yield_grouped) %>%
  .[drop_as_outliers == 1, ] %>%
  .[, .(yield_vol)] %>%
  .[, type := "Dropped"]

rbind(dt_before, dt_after, dt_dropped) %>%
  .[, type := factor(type, levels = c("Before", "After", "Dropped"))] %>%
  ggplot(data = .) +
  geom_histogram(
    aes(x = yield_vol),
    fill = "blue",
    col = "black",
    binwidth = 5,
    alpha = 0.4
  ) +
  facet_grid(type ~ .)
```

## Flow delay correction

```{r eval = F}

get_vg(data = yield_grouped, delay = 0)

get_vg <- function(data, delay) {
  yield_for_shift <-
    data.table(data) %>%
    .[, .(yield_id, yield_vol, ol_yield_vol, ol_speed)]

  yield_shifted_a <-
    data %>%
    mutate(
      mph_to_ms = speed * 1609.34 / (60 * 60),
      travel_time = distance / mph_to_ms,
      cum_travel_time = cumsum(travel_time)
    ) %>%
    data.table() %>%
    .[, .(yield_id, cum_travel_time)]

  yield_shifted_b <-
    copy(yield_shifted_a) %>%
    # === 10 seconds shift ===#
    .[, cum_travel_time := cum_travel_time - delay] %>%
    .[, .(yield_id, cum_travel_time)] %>%
    setnames("yield_id", "yield_id_b")

  id_to_id <- yield_shifted_b[yield_shifted_a, on = "cum_travel_time", roll = "nearest"]

  yield_only_geometry <-
    dplyr::select(data, yield_id) %>%
    rename(yield_id_b = yield_id) %>%
    data.table()

  shifted_polygons <-
    yield_only_geometry[id_to_id, on = "yield_id_b"] %>%
    unique(by = "yield_id_b") %>%
    .[, .(yield_id, geometry)] %>%
    yield_for_shift[., on = "yield_id"] %>%
    .[ol_yield_vol == 0 & ol_speed == 0, ] %>%
    st_as_sf()

  temp_vg <-
    variogram(
      yield_vol ~ 1,
      data = shifted_polygons,
      width = 10,
      cutoff = 200
    )

  return(temp_vg)
}
```

## Yield data reduction when too dense

<!-- Example: Hord_F98_2020 -->

```{r Calculating-median-distance}
(med_distance <- yield$distance %>% median())
```

The following yield point reduction occurs when the medium distance between consecutive points is less than 1 meter.

```{r Reducing-to-smaller-number-of-points-if-dense}
#--- number of observations per reduced group ---#
# about 2 meter (many other yield data points have 2-meter gaps)

if (med_distance < 1) {
  num_obs_rg <- floor(2 / (med_distance))

  yield_reduced_v <- reduce_points_v(
    data_sf = yield_grouped,
    nobs_per_group = num_obs_rg,
    var_interest = "yield_vol",
    ol_var_name = "drop_as_outliers",
    by_var = "group"
  )
} else {
  yield_reduced_v <- data.table(yield_grouped)
}
```

```{r }
cat(
  paste0("Before: ", nrow(yield_grouped), " observations")
)

cat(
  paste0("After: ", nrow(yield_reduced_v), " observations")
)
```


## Create polygons

```{r Creating-polygons-around-points-by-group}
#--- prepare parallel computation ---#
plan(multiprocess, workers = detectCores() - 2)

#--- create polygons ---#
all_polygons <-
  future_lapply(
    #--- list of groups to loop over ---#
    yield_reduced_v$group %>% unique(),
    #--- make polygons by group ---#
    function(x) {
      make_polygons(
        filter(yield_reduced_v, group == x)
      )
    }
  ) %>%
  reduce(c) %>%
  st_as_sfc()

#--- combine with the data ---#
yield_sf <-
  yield_reduced_v %>%
  mutate(geometry = all_polygons) %>%
  st_as_sf() %>%
  st_set_crs(st_crs(yield)) %>%
  #--- get rid of transition polygons by area ---#
  mutate(area = as.numeric(st_area(.))) %>%
  filter(area <= median(area) * 3) %>%
  mutate(yield_id = 1:nrow(.))
```

**Yield map before dropping outliers**
```{r }
tm_shape(yield_sf) +
  tm_polygons(col = "yield_vol", lwd = 0.05) +
  tm_layout_to_add
```

**Yield map after dropping outliers**
```{r }
tm_shape(filter(yield_sf, drop_as_outliers == 0)) +
  tm_polygons(col = "yield_vol", lwd = 0.05) +
  tm_layout_to_add

g_y_poly <-
  ggplot(filter(yield_sf, drop_as_outliers == 0 | yield_vol < 300)) +
  geom_sf(aes(fill = yield_vol * 67.25)) +
  scale_fill_viridis_c(name = "Corn Yield (kg/ha)") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(1, "cm")
  )

ggsave(here("Misc/g_y_poly.png"), g_y_poly, height = 4, width = 7, dpi = 600)
```

