---
title: "Merge Yield and Input Data for DBAT_1_2021"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: true
    number_sections: true
params: 
  topography: false
  ssurgo: false
---

<style type="text/css">

body{ /* Normal  */
      font-size: 20px;
  }
td {  /* Table  */
  font-size: 16px;
}
h1.title {
  font-size: 56px;
}
h1 { /* Header 1 */
  font-size: 48px;
}
h2 { /* Header 2 */
    font-size: 36px;
}
h3 { /* Header 3 */
  font-size: 24px;
}
code.r{ /* Code block */
    font-size: 16px;
}
</style>

```{r setup, cache = F, echo = F}
library(knitr)
knitr::opts_chunk$set(
  cache = FALSE,
  echo = TRUE,
  error = TRUE,
  warning = FALSE,
  cache.lazy = FALSE,
  message = FALSE,
  fig.width = 12,
  fig.height = 12
)
```

```{r packages, cache = FALSE}
# === packages ===#
library(raster)
library(sf)
library(here)
library(ggplot2)
library(tmap)
library(patchwork)
library(measurements)
library(scam)
library(stars)
library(parallel)
library(future.apply)
library(dplyr)
library(tidyverse)
library(modelsummary)
library(jsonlite)
library(data.table)
library(exactextractr)
library(future.apply)
```

```{r preparations, cache = F}
source(
  "https://github.com/tmieno2/OnFarmExperiments/blob/master/Functions/prepare.R?raw=TRUE",
  local = TRUE
)

#--- github ---#
source("https://raw.githubusercontent.com/brittanikedge/DIFM/main/Functions.R")

#--- define field-year ---#
ffy <- "DBAT_1_2021"

#--- read the variable name dictionary ---#
dictionary <-
  fromJSON(
    file.path(
      here("Data", "CommonData"),
      "variable_name_dictionary.json"
    ),
    flatten = TRUE
  ) %>%
  data.table()

#--- define field parameters ---#
trial_info <- get_trial_parameter(ffy)
input_trial_data <- trial_info$input_data_trial

#--- add number of paths in a Rx polygon ---#
#* this is used when filling zero-rate Rx polygons with
#* manufactured input data points

num_paths <- num_paths_here
input_trial_data$num_paths <- num_paths

#---  ---#
plan(multiprocess, workers = detectCores() - 2)
```

```{r set-root-dir}
opts_knit$set(root.dir = here("Data", "Growers", ffy))
```

```{r map-layout, cache = TRUE}
tm_layout_to_add <-
  tm_layout(
    legend.outside = "TRUE",
    frame = FALSE,
    legend.title.size = 2,
    legend.text.size = 1.5
  )
```



# Combine yield and input datasets

## Preparation

### Define some parameters

```{r }
overlap_acceptance_pct <- 0.1
max_dev_ls <- c(10, 10)
ignore_overlap_threshold <- 0.05
```

### Read datasets

**Yield**

```{r }
#--- yield ---#
yield_polygons <- here("Data/Growers", ffy, "Intermediate/yield_polygons.rds") %>%
  readRDS() %>%
  mutate(yield_area = as.numeric(st_area(.)))

# === vis ===#
tm_shape(filter(yield_polygons, drop_as_outliers == 0)) +
  tm_fill(col = "yield_vol", style = "order")
```

**Inputs and their Rx (if exist)**

```{r }
if (length(max_dev_ls) != nrow(input_trial_data)) {
  print("Warning: the number of elements in max_dev_ls does not match with the number of inputs in input_trial_data!")
}
```

```{r }
# input_polygons$rx_polygons

input_polygons <-
  input_trial_data %>%
  mutate(max_dev_allowed = max_dev_ls) %>%
  rowwise() %>%
  mutate(
    input_file_name = here(
      "Data/Growers", ffy, "Intermediate",
      paste0("aa-", input_type, "-", form, "-", input_id, ".rds")
    )
  ) %>%
  # === input ===#
  mutate(input_polygons = list(readRDS(input_file_name))) %>%
  #--- add date ---#
  mutate(input_polygons = list(
    mutate(input_polygons, date = date)
  )) %>%
  # === rx ===#
  mutate(rx_polygons = list(
    if (Rx_data != "none") {
      st_read(here("Data/Growers", ffy, "Raw", Rx_data)) %>%
        mutate(rx_id = 1:nrow(.))
    } else {
      NULL
    }
  ))
```

### Input polygons

```{r }

input_polygons %>%
  mutate(
    tm_map = list(
      tm_shape(input_polygons) +
        tm_fill(
          col = "input_rate",
          title = paste0(input_type, "-", form)
        ) +
        tm_layout_to_add
    )
  ) %>%
  .$tm_map
```

### Rx polygons

```{r eval = F}
input_polygons %>%
  filter(Rx_data != "none") %>%
  mutate(
    tm_map = list(
      tm_shape(rx_polygons) +
        tm_polygons(
          col = form,
          title = paste0(input_type, "-", form)
        ) +
        tm_layout_to_add
    )
  ) %>%
  .$tm_map
```

**Base input rate**

```{r }
n_base <- trial_info$n_base_rate
```

## Data alignment checks

Visually check whether yield data points are aligned well with as-applied input data points.

### Yield and inputs

```{r }
tm_alignment <-
  input_polygons %>%
  mutate(
    tm_map = list(
      tm_shape(input_polygons) +
        tm_borders(lwd = 0.2) +
        tm_shape(st_centroid(yield_polygons)) +
        tm_dots(col = "blue", title = "Yield Points")
    )
  )

tm_alignment$tm_map
```

## Flag unreliable polygons

### Check the overlap and find area-weighted input rates

```{r }

# input_pct$yield_input_int
# input_pct$input_pct_data[[1]][, wm_input_rate_ignore_small]

input_pct <-
  input_polygons %>%
  # === intersect yield and input polygons to check overlaps ===#
  # Every possible cropping intersetcions of yield and nitrogen polygons are made
  mutate(yield_input_int = list(
    st_intersection(
      dplyr::select(yield_polygons, yield_id, yield_area),
      dplyr::select(input_polygons, input_rate, ol_input_rate)
    ) %>%
      # === percentage of overlap relative to the are of a yield polygon ===#
      mutate(sub_pct = as.numeric(st_area(.)) / yield_area) %>%
      data.table() %>%
      #* === flag overlapping input rate observations if below the threshold ===*#
      # flag the interesections that are very small and potentially want to ignore in calculating N for each yield polygon
      .[, ignorable_overlap := fifelse(sub_pct < ignore_overlap_threshold, TRUE, FALSE)]
  )) %>%
  mutate(input_pct_data = list(
    copy(yield_input_int) %>%
      #--- total sub_pct by yield polygon ---#
      .[, tot_sub_pct := sum(sub_pct), by = yield_id] %>%
      #--- calculate sub_pct-weighted MEAN of applied rate ---#
      #* NOT divided by tot_sub_pct
      .[, wm_input_rate := sum(sub_pct * input_rate), by = yield_id] %>%
      #* if total area coverage is less than 5 %, then set the rate to be 0
      .[tot_sub_pct < 0.05, wm_input_rate := 0] %>%
      #--- weighted deviation from the mean ---#
      .[, dev_input_rate := sum(abs(sub_pct * (input_rate - wm_input_rate) / tot_sub_pct)), by = yield_id] %>%
      #--- calculate sub_pct-weighted MEAN of outlier flag ---#
      .[, ol := sum(sub_pct * ol_input_rate) / tot_sub_pct, by = yield_id] %>%
      #--- indicate whether  ---#
      .[ignorable_overlap != TRUE, wm_input_rate_ignore_small := sum(sub_pct * input_rate), by = yield_id] %>%
      #--- order by yield_id ---#
      .[order(yield_id), ] %>%
      .[, .(
        yield_id,
        tot_sub_pct,
        wm_input_rate,
        wm_input_rate_ignore_small,
        dev_input_rate,
        date,
        ol
      )] %>%
      setnames(
        c("wm_input_rate", "wm_input_rate_ignore_small"),
        c("input_rate", "input_rate_ignore_small")
      ) %>%
      unique(by = "yield_id")
  )) %>%
  dplyr::select(input_type, form, var_name_prefix, yield_input_int, input_pct_data, max_dev_allowed) %>%
  mutate(
    # === visualize overlap ===#
    gg_overlap = list(
      ggplot(input_pct_data) +
        geom_histogram(aes(x = tot_sub_pct)) +
        ggtitle(
          paste(
            "Histogram of the percentage of yield polygons \n overlapping with",
            input_type,
            "polygons"
          )
        )
    ),
    # === visualize mix of rates ===#
    gg_dev = list(
      ggplot(input_pct_data) +
        geom_histogram(aes(x = dev_input_rate)) +
        ggtitle(
          paste(
            "Histogram of the deviation of ",
            input_type,
            "\n to its mean by yield_id"
          )
        )
    )
  )
```

```{r }
input_pct$gg_overlap

input_pct$gg_dev
```

### An example polygon with a problem

```{r }
temp_yield_id <- input_pct$input_pct_data[[1]][tot_sub_pct > 2, yield_id][1]

overlapping_input <- input_pct$yield_input_int[[1]][yield_id == temp_yield_id]

temp_yield_poly <- filter(yield_polygons, yield_id == temp_yield_id)

ggplot() +
  geom_sf(data = st_as_sf(overlapping_input), aes(fill = factor(input_rate)), alpha = 0.5) +
  geom_sf(data = temp_yield_poly, color = "red", fill = NA, size = 1)
```

### Flag unreliable observations

```{r }
# === yield data after removing unreliable inputs ===#
yield_input_data_pre <-
  input_pct %>%
  dplyr::select(input_type, var_name_prefix, form, input_pct_data, max_dev_allowed) %>%
  mutate(input_pct_data_pre = list(
    data.table::copy(input_pct_data) %>%
      # === 0 means within the threshold ===#
      #* if tot_sub_pct < 0.05, that can be considered GPS error
      .[, over_lap_test := fifelse(
        (tot_sub_pct > (1 - overlap_acceptance_pct) &
          tot_sub_pct < (1 + overlap_acceptance_pct)) |
          tot_sub_pct < 0.05,
        0,
        1
      )] %>%
      # === 0 means okay ===#
      .[, deviation_test := fifelse(dev_input_rate < max_dev_allowed, 0, 1)] %>%
      .[, .(yield_id, input_rate, input_rate_ignore_small, ol, over_lap_test, deviation_test, date)]
  )) %>%
  mutate(input_pct_data_out = list(
    copy(input_pct_data_pre) %>%
      setnames(
        "date",
        paste0(tolower(input_type), "_", form, "_date")
      ) %>%
      setnames(
        "input_rate",
        paste0(tolower(input_type), "_rate", "_", form)
      ) %>%
      setnames(
        "input_rate_ignore_small",
        paste0(tolower(input_type), "_rate", "_", form, "ignore_small")
      ) %>%
      setnames(
        "ol",
        paste0("ol_", tolower(input_type), "_", form)
      ) %>%
      setnames(
        c("over_lap_test", "deviation_test"),
        c(
          paste0(tolower(input_type), "_overlap", form),
          paste0(tolower(input_type), "_deviation", form)
        )
      ) %>%
      rename_at(vars(-one_of("yield_id")), ~ paste0(ifelse(is.na(var_name_prefix), "", paste0(var_name_prefix, "_")), .))
  ))
```


## Merge with Commercial Rx

This code assumes that the prescription rate variable is named the same as the `form` in the `metadata.json` file.

```{r eval = F}
cm_rx_data <-
  trial_info$rx_data %>%
  rowwise() %>%
  mutate(cm_rx = list(
    st_read(here("Data/Growers", ffy, "Raw", file))
  )) %>%
  mutate(yield_cm_rx = list(
    st_intersection(yield_point, st_transform(cm_rx, st_crs(yield_point))) %>%
      data.table() %>%
      .[, c("yield_id", form), with = FALSE] %>%
      setnames(form, paste0(form, "_", model))
  )) %>%
  pull(yield_cm_rx) %>%
  reduce(left_join, by = "yield_id")
```

## Prepare datasets for analysis and check

### Data for analysis

```{r }
yield_input_data <-
  yield_input_data_pre %>%
  pull(input_pct_data_out) %>%
  reduce(full_join, by = "yield_id") %>%
  inner_join(yield_polygons, ., by = "yield_id") %>%
  # === add base n-rate column ===#
  mutate(n_base_total = trial_info$n_base_data$n_total) %>%
  cbind(., trial_info$n_base_data$n_indiv)
```

### Detailed data for inspection

```{r }
yield_trd <-
  readRDS(
    here(
      "Data/Growers",
      ffy,
      "Intermediate/yield_trd.rds"
    )
  )

dt_trd_int <-
  tibble(
    type = "trial-design",
    int_info = list(yield_trd)
  )

dt_input_int <-
  input_pct %>%
  mutate(type = paste0(input_type, "-", form)) %>%
  dplyr::select(type, yield_input_int) %>%
  rename(int_info = yield_input_int)

detailed_int_data <-
  rbind(dt_trd_int, dt_input_int) %>%
  # === add base n-rate column ===#
  mutate(n_base = n_base)
```

## Visualization

```{r }
viz <-
  yield_input_data_pre %>%
  mutate(input_data_viz = list(
    inner_join(yield_polygons, input_pct_data_pre, by = "yield_id") %>%
      st_as_sf()
  )) %>%
  dplyr::select(input_type, form, input_data_viz) %>%
  mutate(viz_overlap = list(
    ggplot(data = input_data_viz) +
      geom_sf(aes(fill = factor(over_lap_test)), size = 0.4) +
      scale_fill_discrete(name = form)
  )) %>%
  mutate(viz_deviation = list(
    ggplot(data = input_data_viz) +
      geom_sf(aes(fill = factor(deviation_test)), size = 0.4) +
      scale_fill_discrete(name = form)
  ))
```

**Overlap (1 means passing the overlap test)**

```{r }
viz$viz_overlap
```

**Deviation (1 means passing the deviation test)**

```{r }
viz$viz_deviation[[1]]
```

## Save

```{r }
saveRDS(yield_input_data, file = here("Data", "Growers", ffy, "Intermediate/yield_input_data.rds"))

# yield_input_data <- readRDS(file = here("Data", "Growers", ffy, "Intermediate/yield_input_data.rds"))

saveRDS(detailed_int_data, file = here("Data", "Growers", ffy, "DataProcessingReport/detailed_intersection_info.rds"))
```

# Visual Diagnostics

```{r }
trd_polygons <-
  here(
    "Data/Growers", ffy, "TrialDesign",
    paste0(trial_info$tr_design_data, ".shp")
  ) %>%
  st_read() %>%
  mutate(treat_id = 1:nrow(.)) %>%
  dplyr::rename_at(vars(-one_of("geometry")), ~ paste0("trd_", .))

if (!("trd_PlotNum" %in% names(trd_polygons))) {
  trd_polygons <- mutate(trd_polygons, trd_PlotNum = 1:nrow(trd_polygons))
}
```

```{r }
g_diagnostics <-
  yield_input_data_pre %>%
  mutate(plot_data = list(
    left_join(yield_polygons, input_pct_data_pre, by = "yield_id") %>%
      mutate(dropped = (ol != 0 | over_lap_test == 1 | deviation_test == 1))
  )) %>%
  dplyr::select(var_name_prefix, form, plot_data) %>%
  #* === input rate ===*#
  mutate(g_input = list(
    ggplot(plot_data) +
      geom_sf(aes(fill = input_rate), color = NA) +
      geom_sf(data = trd_polygons, color = "red", fill = NA, size = 0.6) +
      scale_fill_distiller(palette = "Spectral", name = "Input Rate") +
      theme_void() +
      ggtitle(paste0(var_name_prefix, "_", form, "Input Rate"))
  )) %>%
  #* === outlier test ===*#
  mutate(g_outlier = list(
    ggplot(plot_data) +
      geom_sf(aes(fill = ol), color = NA) +
      geom_sf(data = trd_polygons, color = "red", fill = NA, size = 0.6) +
      scale_fill_distiller(palette = "Spectral", name = "Outlier") +
      theme_void() +
      ggtitle(paste0(var_name_prefix, "_", form, "Outlier"))
  )) %>%
  #* === overlap test ===*#
  mutate(g_overlap = list(
    ggplot(plot_data) +
      geom_sf(aes(fill = factor(over_lap_test)), color = NA) +
      geom_sf(data = trd_polygons, color = "darkgreen", fill = NA, size = 0.6) +
      scale_fill_discrete(name = "overlap") +
      theme_void() +
      ggtitle(paste0(var_name_prefix, "_", form, "Overlap"))
  )) %>%
  #* === deviation test ===*#
  mutate(g_dev = list(
    ggplot(plot_data) +
      geom_sf(aes(fill = factor(deviation_test)), color = NA) +
      geom_sf(data = trd_polygons, color = "darkgreen", fill = NA, size = 0.6) +
      scale_fill_discrete(name = "Deviation") +
      theme_void() +
      ggtitle(paste0(var_name_prefix, "_", form, "Deviation"))
  )) %>%
  #* === drop or not ===*#
  mutate(g_dropped = list(
    ggplot(plot_data) +
      geom_sf(aes(fill = factor(dropped)), color = NA) +
      geom_sf(data = trd_polygons, color = "darkgreen", fill = NA, size = 0.6) +
      scale_fill_discrete(name = "Dropped") +
      theme_void() +
      ggtitle(paste0(var_name_prefix, "_", form, "Drop or not (1 means drop)"))
  ))
```


```{r }
i <- 1
for (i in 1:nrow(g_diagnostics)) {
  temp <- g_diagnostics[i, ]
  print(temp$g_input[[1]])
  print(temp$g_outlier[[1]])
  print(temp$g_overlap[[1]])
  print(temp$g_dev[[1]])
  print(temp$g_dropped[[1]])
}
```

# Figure created here!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

```{r}
data.table(yield_input_data)[FA1_n_deviationUAN32 == 0, ]

temp_yield_id <- data.table(yield_input_data)[FA1_n_deviationUAN32 == 1, median(yield_id)]

# data.table(yield_input_data)[, mean(FA1_n_overlapUAN32, na.rm = TRUE)]

temp_polygon <- filter(yield_input_data, yield_id == temp_yield_id) %>%
  st_buffer(dist = 30)

yield_temp <- yield_input_data[temp_polygon, ]

yield_input_data[temp_polygon, ]

input_temp <- input_polygons[1, ]$input_polygons[[1]][temp_polygon, ]

g_issues <-
  ggplot() +
  geom_sf(data = input_temp, aes(fill = input_rate * 1.12085), color = NA) +
  geom_sf(data = yield_temp, fill = NA, aes(color = "Yield Polygon"), size = 1) +
  scale_color_discrete(name = "") +
  scale_fill_viridis_c(name = "Nitrogen Rate (kg/ha)") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(1, "cm")
  )

ggsave(here("Misc/g_issues.png"), g_issues, width = 7, height = 4, dpi = 600)
```


## lack of overlap 

```{r}
data.table(yield_input_data)[FA1_n_deviationUAN32 == 0, ]

temp_yield_id <- data.table(yield_input_data)[FA1_n_deviationUAN32 == 1, median(yield_id)]

temp_polygon <- filter(yield_input_data, yield_id == temp_yield_id) %>%
  st_buffer(dist = 30)

yield_temp <- yield_input_data[temp_polygon, ]

yield_input_data[temp_polygon, ]

input_temp <- input_polygons[1, ]$input_polygons[[1]][temp_polygon, ]

g_issues <-
  ggplot() +
  geom_sf(data = input_temp, aes(fill = input_rate * 1.12085), color = NA) +
  geom_sf(data = yield_temp, fill = NA, color = "red", size = 0.8) +
  scale_fill_viridis_c(name = "Nitrogen Rate (kg/ha)") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(1, "cm")
  )

ggsave(here("Misc/g_issues.png"), g_issues, width = 7, height = 4, dpi = 600)
```